---
title: "Demographics analysis 3 - HSCP and COVID"
subtitle: "COVID Admissions - January 2020 to Febraury 2022"
output:
  html_document:
    df_print: paged
---
HSCP - 9-digit code for Health and Social Care Partnerships (2016) of residence

HB - 9-digit code for health board of treatment based on boundaries as at 1st 
April 2019

```{r}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(ggplot2)
library(lubridate)
library(stringr)
```
<br> 

Read In Data

```{r}
covid_admissions_HSCP_agesex <- read_csv(here("../../raw_data/covid_data/hospital_admissions_hscp_agesex_20220302.csv"))

covid_admissions_HSCP_simd <- read_csv(here("../../raw_data/covid_data/hospital_admissions_hscp_simd_20220302.csv"))

```

**Clean Names**

```{r}
covid_admissions_HSCP_agesex <- janitor::clean_names(covid_admissions_HSCP_agesex)
covid_admissions_HSCP_simd <- janitor::clean_names(covid_admissions_HSCP_simd)
```

**Change week_ending column to a date and creating new day, month, year AND ** 
**a month and year col**
```{r}
#HSCP x age sex
covid_admissions_HSCP_agesex <- covid_admissions_HSCP_agesex %>%
  mutate(week_ending = ymd(week_ending))

covid_admissions_HSCP_agesex <- covid_admissions_HSCP_agesex %>%
  mutate(
    wk_ending_day = day(week_ending),
    wk_ending_month = month(week_ending, label = TRUE, abbr = FALSE),
    wk_ending_year = year(week_ending)
  ) %>% 
  unite(wk_ending_yr_month, 
        wk_ending_month, wk_ending_year, 
        remove = FALSE, 
        sep = " ")
covid_admissions_HSCP_agesex

#HSCP x simd
covid_admissions_HSCP_simd <- covid_admissions_HSCP_simd %>%
  mutate(week_ending = ymd(week_ending))

covid_admissions_HSCP_simd <- covid_admissions_HSCP_simd %>%
  mutate(
    wk_ending_day = day(week_ending),
    wk_ending_month = month(week_ending, label = TRUE, abbr = FALSE),
    wk_ending_year = year(week_ending)
  ) %>% 
  unite(wk_ending_yr_month, 
        wk_ending_month, wk_ending_year, 
        remove = FALSE, 
        sep = " ")
covid_admissions_HSCP_simd
```


<hr>

<br>

#### What are we working with?
**Looking to see if i can join data? Probably??? All the HSPC corresponding **
**health board values are S08000015 - s08000032 **
**https://www.opendata.nhs.scot/dataset/geography-codes-and-labels/resource/944765d7-d0d9-46a0-b377-abb3de51d08e**

**The HB data set is all S08000015 - s08000032** 

```{r}
covid_admissions_HSCP_agesex %>% 
  distinct(hscp)
```

We only need ACUTE patients: 

admission_type has: All, Emergency and Planned
* all and emergency always have similar figures compared with planned.

```{r}
covid_admissions_HSCP_agesex%>% 
  group_by(admission_type) %>% 
  summarise(total = n())
covid_admissions_HSCP_agesex %>% 
  group_by(age_group) %>% 
  summarise(total = n())

covid_admissions_HSCP_simd%>% 
  group_by(admission_type) %>% 
  summarise(total = n())
covid_admissions_HSCP_simd %>% 
  group_by(simd_quintile) %>% 
  summarise(total = n())
```


* let's remove "all" as we only want acute 
```{r}
covid_admissions_HSCP_agesex <- covid_admissions_HSCP_agesex %>% 
  filter(admission_type == "Emergency")

covid_admissions_HSCP_simd <- covid_admissions_HSCP_simd %>% 
  filter(admission_type == "Emergency")
```

* We can remove simd NAs
* ANd we don't need ALL ages either... edit: yes we do:
 * The "sex" column "Male, Female, All" - m & f only come under "All ages"!
```{r}
covid_admissions_HSCP_simd <- covid_admissions_HSCP_simd %>% 
  drop_na(simd_quintile)

covid_admissions_HSCP_AGE <- covid_admissions_HSCP_agesex %>% 
  filter(age_group != "All ages")
```

<hr>

<br>

## HSCP - sex and age

### HB - Age - Total Covid admissions
<br>

Firstly need a total column for each age group per month:
```{r}

#Under 5
monthly_covid_ads_HSCP_under5 <- covid_admissions_HSCP_AGE %>% 
  filter(age_group == "Under 5") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

monthly_past_average_HSCP_under5 <- monthly_covid_ads_HSCP_under5 %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
monthly_past_average_HSCP_under5

#5 - 14
monthly_covid_ads_HSCP_5to14 <- covid_admissions_HSCP_AGE %>% 
  filter(age_group == "5 - 14") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

monthly_past_average_HSCP_5to14 <- monthly_covid_ads_HSCP_5to14 %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
monthly_past_average_HSCP_5to14
  
#15 - 44
monthly_covid_ads_HSCP_15to44 <- covid_admissions_HSCP_AGE %>% 
  filter(age_group == "15 - 44") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

monthly_past_average_HSCP_15to44 <- monthly_covid_ads_HSCP_15to44 %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
monthly_past_average_HSCP_15to44

# 45 - 64
a <- covid_admissions_HSCP_AGE %>% 
  filter(age_group == "45 - 64") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

a2 <- a %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
a2

#65-74  
b <- covid_admissions_HSCP_AGE %>% 
  filter(age_group == "65 - 74") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

b2 <- b %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
b2

#75-84
c <- covid_admissions_HSCP_AGE %>% 
  filter(age_group == "75 - 84") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

c2 <- c %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
c2

#85 and over
d <- covid_admissions_HSCP_AGE %>% 
  filter(age_group == "85 and over") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

d2 <- d %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
d2
```

bind together so we have totals for each month:
```{r}
covid_admissions_HSCP_age_totals <- bind_rows(monthly_past_average_HSCP_under5, 
                                            monthly_past_average_HSCP_5to14,
                                            monthly_past_average_HSCP_15to44,
                                            a2, b2, c2, d2)
covid_admissions_HSCP_age_totals
```

April 2020 will be shown first and ages are a mess. So:

```{r}
dates <- c("January 2020", "February 2020", "March 2020", "April 2020", 
           "May 2020", "June 2020", "July 2020", "August 2020", 
           "September 2020", "October 2020", "November 2020", "December 2020", 
           "January 2021", "February 2021", "March 2021", "April 2021", 
           "May 2021", "June 2021", "July 2021", "August 2021", 
           "September 2021", "October 2021", "November 2021", "December 2021", 
           "January 2022", "February 2022")

summer_dates <- c("April 2020", "May 2020", "June 2020", "July 2020", 
                  "August 2020", "September 2020", "April 2021", "May 2021", 
                  "June 2021", "July 2021", "August 2021", "September 2021")

winter_dates <- c("January 2020", "February 2020", "March 2020", "October 2020",
                  "November 2020", "December 2020", "January 2021", 
                  "February 2021", "March 2021", "October 2021", 
                  "November 2021", "December 2021", "January 2022", 
                  "February 2022")
```

<br>
**PLOT OF Total COVID admissions per month by age group**

```{r}
covid_admissions_HSCP_age_totals %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month, 
      group = age_group, 
      colour = age_group)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  geom_line()+
  labs(x = "Month and Year",
       y = "Total admissions",
       title = "HSCP Total admissions in COVID times per month by age group",
       subtitle = "January, 2020 - February 2022",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


let's try again to see if there's difference in age groups and seasons:
* Winter = Q4 and Q1 
* Summer = Q2 and Q3
```{r}
highlight_winter_HSCP_age <- covid_admissions_HSCP_age_totals %>% 
  filter(str_detect(wk_ending_yr_month, 
                    "October|November|December|January|February|March"))

highlight_summer_HSCP_age <- covid_admissions_HSCP_age_totals %>% 
  filter(str_detect(wk_ending_yr_month, 
                    "April|May|June|July|August|September"))
```

Re-trying above graph -  messy but can be looked into.
* Total COVID admissions per month by age group
```{r}
covid_admissions_HSCP_age_totals %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month, 
      group = age_group, 
      fill = age_group)+
  scale_x_discrete(limits = dates) +
  geom_col()+
  facet_wrap(~age_group)+
  labs(x = "Month",
       y = "Total admissions",
       title = "HSCP Total admissions in COVID times per month by age group",
       subtitle = "January, 2020 - February 2022",
       fill = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


**THE ABOVE BUT WINTER vs SUMMER**

* Total COVID admissions per month by age group
* SUMMER
```{r}

highlight_summer_HSCP_age %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month, 
      group = age_group, 
      fill = age_group)+
  scale_x_discrete(limits = summer_dates) +
  geom_col()+
  facet_wrap(~age_group)+
  labs(x = "Month",
       y = "Total admissions",
       title = "Total admissions in COVID times per SUMMER month, by age group",
       subtitle = "Summer Months: 
       April, 2020 - September 2020 
       and April, 2021 - September 2021",
       fill = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


* Total COVID admissions per month by age group
* WINTER

```{r}

highlight_winter_HSCP_age %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month, 
      group = age_group, 
      fill = age_group)+
  scale_x_discrete(limits = winter_dates) +
  geom_col()+
  facet_wrap(~age_group)+
  labs(x = "Month and Year",
       y = "Total admissions",
       title = "HSCP Total admissions in COVID times per WINTER month by age group",
       subtitle = "Winter Months: 
       January, 2020 - March 2020; 
       October 2020 - March 2021; 
       October 2021 - February 2022",
       fill = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


<hr>
<br>

### HSCP -Age- Weekly COVID admissions against the average admissions in 2018 & 19

* wow this is messy 

```{r}
covid_admissions_HSCP_age_totals %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  group_by(age_group) %>% 
  ggplot()+
  aes(x = number_admissions, 
      y = average20182019, 
      colour = age_group)+
  geom_point() +
  labs(x = "Weekly number of admissions in COVID times",
       y = "Average weekly admissions to hospital in 2018-2019",
       title = "Weekly admissions in COVID TIMES per age group against the 
       equivalent average weekly admissions in previous years",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

**The above again just looking to see if there are differences in age groups**
* Weekly COVID admissions per age group against the average weekly admissions in 
previous years (2018&19)

```{r}
covid_admissions_HSCP_age_totals %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>% 
  group_by(age_group) %>% 
  ggplot()+
  aes(x = number_admissions, 
      y = average20182019, 
      colour = age_group)+
  geom_point() +
  facet_wrap(~age_group) +
  labs(x = "Weekly number of admissions in COVID times",
       y = "Average weekly admissions to hospital in 2018-2019",
       title = "Weekly admissions in COVID TIMES per age group against the 
       equivalent average weekly admissions in previous years",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))

```

**The above but by month instead of EVERY input**
```{r}
covid_admissions_HSCP_age_totals %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  group_by(age_group) %>% 
  ggplot()+
  aes(x = total_admissions_per_month, 
      y = total_past_average_ads_per_month, 
      colour = age_group)+
  geom_point() +
  labs(x = "Monthly number of admissions in COVID times",
       y = "Average monthly admissions to hospital in 2018-2019",
       title = "Monthly admissions in COVID TIMES per age group against the 
       equivalent average monthly admissions in previous years",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

Facet by age group

```{r}
covid_admissions_HSCP_age_totals %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  group_by(age_group) %>% 
  ggplot()+
  aes(x = total_admissions_per_month, 
      y = total_past_average_ads_per_month, 
      colour = age_group)+
  geom_point() +
    facet_wrap(~age_group) +
  labs(x = "Monthly number of admissions in COVID times",
       y = "Average monthly admissions to hospital in 2018-2019",
       title = "Monthly admissions in COVID TIMES per age group against the 
       equivalent average monthly admissions in previous years",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


<hr>

### HSCP - Sex - Average Covid Admissions vs Average 2018 and 2019 admissions
 
 * remember that the agesex df has male, female, and all - male and female come
 under "All ages" but not specific age groups
 
 * Creating columns for monthly totals - admissions, and average past ads
 * FOR FEMALES (males below):

```{r}

monthly_covid_ads_HSCP_agesex_female <- covid_admissions_HSCP_agesex %>% 
  filter(sex == "Female") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

monthly_past_average_HSCP_agesex_female <- monthly_covid_ads_HSCP_agesex_female %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
monthly_past_average_HSCP_agesex_female
  
```


 * Creating columns for monthly totals - admissions, and average past ads
 * FOR MALES (females above):
```{r}
# covid_admissions_HB_agesex %>% 
#   distinct(sex)
# view(covid_admissions_HB_agesex)

monthly_covid_ads_HSCP_agesex_male <- covid_admissions_HSCP_agesex %>% 
  filter(sex == "Male") %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month = sum(number_admissions))

monthly_past_average_HSCP_agesex_male <- monthly_covid_ads_HSCP_agesex_male %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
monthly_past_average_HSCP_agesex_male
  
```


bringing male and females with totals per month together:
```{r}
covid_admissions_HSCP_sex_totals <- bind_rows(monthly_past_average_HSCP_agesex_female, 
                                     monthly_past_average_HSCP_agesex_male)
covid_admissions_HSCP_sex_totals
```


Total admissions across HSCPs by gender for week ending (this doesn't work):
```{r}
covid_admissions_HSCP_sex_totals%>% 
  group_by(sex) %>% 
  ggplot()+
  aes(x = total_admissions_per_month, 
      y = total_past_average_ads_per_month,
      colour = sex)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  geom_line()+
  labs(title = "I AM AWARE THIS IS AWFUL")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

Trying again - Total admissions across HSCPs by gender for week ending 
```{r}
covid_admissions_HSCP_sex_totals %>% 
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month,
      fill = sex)+
  scale_x_discrete(limits = dates) +
  geom_col(position = "dodge")+
  labs(x = "Month and Year",
       y = "Total Admissions per Month",
       title = "HSCP Total Admissions in COVID times per Month by Sex", 
       fill = "Sex")+  
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


weekly  
```{r}
covid_admissions_HSCP_sex_totals %>% 
  group_by(sex) %>% 
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month,
      colour = sex)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  geom_line()+ # where is the line?
  labs(x = "Month and Year",
       y = "Total Admissions",
       title = "HSCP Monthly admissions in COVID times by sex",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "Sex") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


<hr> 

## HEALTH BOARD - SIMD 

### HSCP - SIMD - Total Admissions 

```{r}
covid_admissions_HSCP_simd %>% 
  group_by(wk_ending_yr_month, simd_quintile) %>% 
  ggplot()+
  aes(x = number_admissions, 
      y = average20182019,
      colour = simd_quintile)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  geom_line()+
  labs(title = "I AM AWARE THIS IS AWFUL")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
  
```

So, creating monthly total columns for each level of simd:
```{r}
simd1 <- covid_admissions_HSCP_simd %>% 
  filter(simd_quintile == "1") %>%
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month_simd = sum(number_admissions))

simd1 <- simd1 %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
simd1

#2
simd2 <- covid_admissions_HSCP_simd %>% 
  filter(simd_quintile == "2") %>%
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month_simd = sum(number_admissions))

simd2 <- simd2 %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
simd2

#3
simd3 <- covid_admissions_HSCP_simd %>% 
  filter(simd_quintile == "3") %>%
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month_simd = sum(number_admissions))

simd3 <- simd3 %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
simd3

#4
simd4 <- covid_admissions_HSCP_simd %>% 
  filter(simd_quintile == "4") %>%
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month_simd = sum(number_admissions))

simd4 <- simd4 %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
simd4

#5
simd5 <- covid_admissions_HSCP_simd %>% 
  filter(simd_quintile == "5") %>%
  group_by(wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month_simd = sum(number_admissions))

simd5 <- simd5 %>% 
  group_by(wk_ending_yr_month) %>% 
  mutate(total_past_average_ads_per_month = sum(average20182019))
simd5
  
```

bind these together:
```{r}
covid_admissions_HSCP_simd_totals <- bind_rows(simd1, simd2, simd3, simd4, simd5)
covid_admissions_HSCP_simd_totals
```


**PLOT total admissions per simd per month**
```{r}
covid_admissions_HSCP_simd_totals %>% 
   mutate(simd_quintile = fct_relevel(as.character(simd_quintile, 
                                  "1", "2", "3", "4", "5"))) %>%
  group_by(wk_ending_yr_month, simd_quintile) %>% 
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month_simd, 
      group = simd_quintile, 
      colour = simd_quintile)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  geom_line()+
  labs(x = "Month and year",
       y = "Total admissions",
       title = "HSCP Total admissions in COVID times per month by SIMD Level",
       subtitle = "January, 2020 - February 2022",
       colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

SIMD - AVERAGE admissions 2018 and 2019 against COVID admissions 
```{r}
covid_admissions_HSCP_simd_totals %>% 
   mutate(simd_quintile = fct_relevel(as.character(simd_quintile, 
                                  "1", "2", "3", "4", "5"))) %>%
  group_by(simd_quintile) %>% 
  ggplot()+
  aes(x = total_admissions_per_month_simd, 
      y = total_past_average_ads_per_month,
      colour = simd_quintile)+
  geom_point()+
  labs(x = "Monthly number of admissions",
       y = "Average monthly admissions to hospital in 2018-2019",
       title = "Monthly admissions in COVID Times per SIMD level against the 
       equivalent average monthly admissions in previous years",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```  


