---
title: "Demographics analysis 2 - HB and COVID"
subtitle: "COVID Admissions - January 2020 to Febraury 2022"
output:
  html_document:
    df_print: paged
---


not needed


summer_dates <- c("April 2020", "May 2020", "June 2020", "July 2020", 
                  "August 2020", "September 2020", "April 2021", "May 2021", 
                  "June 2021", "July 2021", "August 2021", "September 2021")

winter_dates <- c("January 2020", "February 2020", "March 2020", "October 2020",
                  "November 2020", "December 2020", "January 2021", 
                  "February 2021", "March 2021", "October 2021", 
                  "November 2021", "December 2021", "January 2022", 
                  "February 2022")





HB - 9-digit code for health board of treatment based on boundaries as at 1st 
April 2019

```{r}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(ggplot2)
library(lubridate)
library(stringr)
```
<br> 

Read In Data

```{r}

covid_admissions_HB_simd <- read_csv(here("../../raw_data/covid_data/hospital_admissions_hb_simd_20220302.csv"))

```

**Clean Names**

```{r}
covid_admissions_HB_agesex <- janitor::clean_names(covid_admissions_HB_agesex)
covid_admissions_HB_simd <- janitor::clean_names(covid_admissions_HB_simd)
```

**Change week_ending column to a date and creating new month, year AND ** 
**a month and year col** 
```{r}
# Health Board x Age sex
covid_admissions_HB_agesex <- covid_admissions_HB_agesex %>%
  mutate(week_ending = ymd(week_ending))

covid_admissions_HB_agesex <- covid_admissions_HB_agesex %>%
  mutate(
    wk_ending_month = month(week_ending, label = TRUE, abbr = FALSE),
    wk_ending_year = year(week_ending)
  ) %>% 
  unite(wk_ending_yr_month, 
        wk_ending_month, wk_ending_year, 
        remove = FALSE, 
        sep = " ")
covid_admissions_HB_agesex


#Health board x simd
covid_admissions_HB_simd <- covid_admissions_HB_simd %>%
  mutate(week_ending = ymd(week_ending))

covid_admissions_HB_simd <- covid_admissions_HB_simd %>%
  mutate(
    wk_ending_month = month(week_ending, label = TRUE, abbr = FALSE),
    wk_ending_year = year(week_ending)
  ) %>% 
  unite(wk_ending_yr_month, 
        wk_ending_month, wk_ending_year, 
        remove = FALSE, 
        sep = " ")
covid_admissions_HB_simd

```
<br>

Adding a quarter and year column to both
```{r}
covid_admissions_HB_agesex <- covid_admissions_HB_agesex %>% 
  mutate(year_quarter = case_when(
    wk_ending_yr_month == "January 2020" ~ "2020Q1",
    wk_ending_yr_month == "February 2020" ~ "2020Q1",
    wk_ending_yr_month == "March 2020" ~ "2020Q1",
    wk_ending_yr_month == "April 2020" ~ "2020Q2",
    wk_ending_yr_month == "May 2020" ~ "2020Q2",
    wk_ending_yr_month == "June 2020" ~ "2020Q2",
    wk_ending_yr_month == "July 2020" ~ "2020Q3",
    wk_ending_yr_month == "August 2020" ~ "2020Q3",
    wk_ending_yr_month == "September 2020" ~ "2020Q3",
    wk_ending_yr_month == "October 2020" ~ "2020Q4",
    wk_ending_yr_month == "November 2020" ~ "2020Q4",
    wk_ending_yr_month == "December 2020" ~ "2020Q4",
    wk_ending_yr_month == "January 2021" ~ "2021Q1",
    wk_ending_yr_month == "February 2021" ~ "2021Q1",
    wk_ending_yr_month == "March 2021" ~ "2021Q1",
    wk_ending_yr_month == "April 2021" ~ "2021Q2",
    wk_ending_yr_month == "May 2021" ~ "2021Q2",
    wk_ending_yr_month == "June 2021" ~ "2021Q2",
    wk_ending_yr_month == "July 2021" ~ "2021Q3",
    wk_ending_yr_month == "August 2021" ~ "2021Q3",
    wk_ending_yr_month == "September 2021" ~ "2021Q3",
    wk_ending_yr_month == "October 2021" ~ "2021Q4",
    wk_ending_yr_month == "November 2021" ~ "2021Q4",
    wk_ending_yr_month == "December 2021" ~ "2021Q4",
    wk_ending_yr_month == "January 2022" ~ "2022Q1",
    wk_ending_yr_month == "February 2022" ~ "2022Q1",
  ))

covid_admissions_HB_simd <- covid_admissions_HB_simd %>% 
  mutate(year_quarter = case_when(
    wk_ending_yr_month == "January 2020" ~ "2020Q1",
    wk_ending_yr_month == "February 2020" ~ "2020Q1",
    wk_ending_yr_month == "March 2020" ~ "2020Q1",
    wk_ending_yr_month == "April 2020" ~ "2020Q2",
    wk_ending_yr_month == "May 2020" ~ "2020Q2",
    wk_ending_yr_month == "June 2020" ~ "2020Q2",
    wk_ending_yr_month == "July 2020" ~ "2020Q3",
    wk_ending_yr_month == "August 2020" ~ "2020Q3",
    wk_ending_yr_month == "September 2020" ~ "2020Q3",
    wk_ending_yr_month == "October 2020" ~ "2020Q4",
    wk_ending_yr_month == "November 2020" ~ "2020Q4",
    wk_ending_yr_month == "December 2020" ~ "2020Q4",
    wk_ending_yr_month == "January 2021" ~ "2021Q1",
    wk_ending_yr_month == "February 2021" ~ "2021Q1",
    wk_ending_yr_month == "March 2021" ~ "2021Q1",
    wk_ending_yr_month == "April 2021" ~ "2021Q2",
    wk_ending_yr_month == "May 2021" ~ "2021Q2",
    wk_ending_yr_month == "June 2021" ~ "2021Q2",
    wk_ending_yr_month == "July 2021" ~ "2021Q3",
    wk_ending_yr_month == "August 2021" ~ "2021Q3",
    wk_ending_yr_month == "September 2021" ~ "2021Q3",
    wk_ending_yr_month == "October 2021" ~ "2021Q4",
    wk_ending_yr_month == "November 2021" ~ "2021Q4",
    wk_ending_yr_month == "December 2021" ~ "2021Q4",
    wk_ending_yr_month == "January 2022" ~ "2022Q1",
    wk_ending_yr_month == "February 2022" ~ "2022Q1",
  ))
```

<br>

<hr>

<br>

#### What are we working with?
**Looking to see if i can join data? Probably??? All the HSPC corresponding **
**health board values are S08000015 - s08000032 **
**https://www.opendata.nhs.scot/dataset/geography-codes-and-labels/resource/944765d7-d0d9-46a0-b377-abb3de51d08e**

**The HB data set is all S08000015 - s08000032** 

```{r}
covid_admissions_HB_agesex %>% 
  distinct(hb)
```

We only need ACUTE patients: 

admission_type has: All, Emergency and Planned
* all and emergency always have similar figures compared with planned.

```{r}
covid_admissions_HB_agesex %>% 
  group_by(admission_type) %>% 
  summarise(total = n())
covid_admissions_HB_agesex %>% 
  group_by(age_group) %>% 
  summarise(total = n())

covid_admissions_HB_simd%>% 
  group_by(admission_type) %>% 
  summarise(total = n())
covid_admissions_HB_simd %>% 
  group_by(simd_quintile) %>% 
  summarise(total = n())
```


* let's remove "all" as we only want acute 
```{r}
covid_admissions_HB_agesex <- covid_admissions_HB_agesex %>% 
  filter(admission_type == "Emergency")

covid_admissions_HB_simd <- covid_admissions_HB_simd %>% 
  filter(admission_type == "Emergency")

```

ANd we don't need ALL ages either... edit: yes we do:
 * The "sex" column "Male, Female, All" - m & f only come under "All ages"!
```{r}
covid_admissions_HB_AGE <- covid_admissions_HB_agesex %>% 
  filter(age_group != "All ages")
```

<hr>

<br>

## HEALTH BOARD - sex and age

### HB - Age - Total Covid admissions
<br>

Firstly need a total column for each age group per month and for equivalent 
average18/19 - MONTH AND QUARTER:
```{r}
covid_admissions_HB_AGE <- covid_admissions_HB_AGE %>% 
  group_by(age_group, wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month_age = sum(number_admissions))

covid_admissions_HB_AGE <- covid_admissions_HB_AGE %>% 
  group_by(age_group, wk_ending_yr_month) %>% 
  mutate(admissions_2018_19_per_month_age = sum(average20182019))
covid_admissions_HB_AGE

covid_admissions_HB_AGE <- covid_admissions_HB_AGE %>% 
  group_by(age_group, year_quarter) %>% 
  mutate(total_admissions_per_quarter_age = sum(number_admissions))
covid_admissions_HB_AGE

covid_admissions_HB_AGE <- covid_admissions_HB_AGE %>%
  group_by(age_group, year_quarter) %>% 
  mutate(admissions_2018_19_per_quarter_age = sum(average20182019))
covid_admissions_HB_AGE

```


April 2020 will be shown first and ages are a mess. So:

```{r}
dates <- c("January 2020", "February 2020", "March 2020", "April 2020", 
           "May 2020", "June 2020", "July 2020", "August 2020", 
           "September 2020", "October 2020", "November 2020", "December 2020", 
           "January 2021", "February 2021", "March 2021", "April 2021", 
           "May 2021", "June 2021", "July 2021", "August 2021", 
           "September 2021", "October 2021", "November 2021", "December 2021", 
           "January 2022", "February 2022")

summer_dates <- c("April 2020", "May 2020", "June 2020", "July 2020", 
                  "August 2020", "September 2020", "April 2021", "May 2021", 
                  "June 2021", "July 2021", "August 2021", "September 2021")

winter_dates <- c("January 2020", "February 2020", "March 2020", "October 2020",
                  "November 2020", "December 2020", "January 2021", 
                  "February 2021", "March 2021", "October 2021", 
                  "November 2021", "December 2021", "January 2022", 
                  "February 2022")

quarter_dates <- c("2016Q3", "2016Q4", "2017Q1", "2017Q2", "2017Q3", "2017Q4", 
                   "2018Q1", "2018Q2", "2018Q3", "2018Q4", "2019Q1", "2019Q2", 
                   "2019Q3", "2019Q4", "2020Q1", "2020Q2", "2020Q3", "2020Q4",
                   "2021Q1", "2021Q2", "2021Q3")

q1 <- c("January 2020", "February 2020", "March 2020", "January 2021", 
        "February 2021", "March 2021", "January 2022", "February 2022")

q2 <- c("April 2020", "May 2020", "June 2020", "April 2021", "May 2021", 
        "June 2021") 

q3 <- c("July 2020", "August 2020", "September 2020", "July 2021", 
        "August 2021", "September 2021")

q4 <- c("October 2020", "November 2020", "December 2020", "October 2021", 
        "November 2021", "December 2021")
```

<br>
**PLOT OF Total COVID admissions per month by age group**


```{r}
covid_admissions_HB_AGE %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month_age, 
      group = age_group, 
      colour = age_group)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  geom_line()+
  labs(x = "Month and Year",
       y = "Total admissions",
       title = "(HB) Total Admissions during COVID, per month and across 
       Scotland by age group",
       subtitle = "January, 2020 - February 2022",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


By Quarter: 
```{r}
covid_admissions_HB_AGE %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  ggplot()+
  aes(x = year_quarter, 
      y = total_admissions_per_quarter_age, 
      group = age_group, 
      colour = age_group)+
  geom_point()+
  geom_line()+
  labs(x = "Yearly Quarters",
       y = "Total admissions",
       title = "(HB) Total Admissions during COVID, per quarter and across 
       Scotland by age group",
       subtitle = "January, 2020 - February 2022",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```





<hr>
<br>

### HB -Age- Weekly COVID admissions against the average admissions in 2018 & 19

```{r}
covid_admissions_HB_AGE %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  group_by(age_group) %>% 
  ggplot()+
  aes(x = number_admissions, 
      y = average20182019, 
      colour = age_group)+
  geom_point() +
  labs(x = "Weekly number of admissions",
       y = "Average weekly admissions in 2018-2019",
       title = "(HB) Weekly admissions across Scotland in COVID times against 
       the equivalent average weekly admissions in previous years 
       (per age group)",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```





**The above but by month instead of EVERY input**
```{r}
covid_admissions_HB_AGE %>% 
  mutate(age_group = fct_relevel(age_group, 
                                 "Under 5", "5 - 14", "15 - 44", 
                                 "45 - 64", "65 - 74", "75 - 84", 
                                 "85 and over")) %>%
  group_by(age_group) %>% 
  ggplot()+
  aes(x = total_admissions_per_month_age, 
      y = admissions_2018_19_per_month_age, 
      colour = age_group)+
  geom_point() +
  labs(x = "Monthly number of admissions",
       y = "Average monthly admissions in 2018-2019",
       title = "(HB) Monthly admissions across Scotland in COVID times against 
       the equivalent average monthly admissions in previous years 
       (per age group)",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

<hr>

### HB - Sex - Average Covid Admissions vs Average 2018 and 2019 admissions
 
 * remember that the agesex df has male, female, and all - male and female come
 under "All ages" but not specific age groups
 
 * Creating columns for monthly and quarterly totals - admissions, and average past ads
```{r}
covid_admissions_HB_SEX <- covid_admissions_HB_agesex %>%
  filter(sex != "All")
# view(covid_admissions_HB_agesex)

covid_admissions_HB_SEX <- covid_admissions_HB_SEX %>% 
  group_by(sex, wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month_sex = sum(number_admissions))

covid_admissions_HB_SEX <- covid_admissions_HB_SEX %>%  
  group_by(sex, wk_ending_yr_month) %>% 
  mutate(admissions_2018_19_per_month_sex = sum(average20182019))

covid_admissions_HB_SEX <- covid_admissions_HB_SEX %>% 
  group_by(sex, year_quarter) %>% 
  mutate(total_admissions_per_quarter_sex = sum(number_admissions))

covid_admissions_HB_SEX <- covid_admissions_HB_SEX %>%  
  group_by(sex, year_quarter) %>% 
  mutate(admissions_2018_19_per_quarter_sex = sum(average20182019))

```


Total admissions across HBs by gender for week ending:

```{r}
covid_admissions_HB_SEX %>% 
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month_sex,
      group = sex, 
      colour = sex)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  geom_line()+
  labs(x = "Month and Year",
       y = "Total Admissions per Month",
       title = "Total Admissions across Health Boards in COVID Times per Month,
       by Sex", 
       fill = "Sex")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

Trying again - Total admissions across HBs by gender for week ending 
```{r}
covid_admissions_HB_SEX %>% 
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month_sex,
      fill = sex)+
  scale_x_discrete(limits = dates) +
  geom_col(position = "dodge")+
  labs(x = "Month and Year",
       y = "Total Admissions per Month",
       title = "Total Admissions across Health Boards in COVID Times per Month,
       by Sex", 
       fill = "Sex")+  
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```


By quarter
```{r}
covid_admissions_HB_SEX %>% 
  ggplot()+
  aes(x = year_quarter, 
      y = total_admissions_per_quarter_sex,
      group = sex, 
      colour = sex)+
  geom_point()+
  geom_line()+
  labs(x = "Quarter and Year",
       y = "Total Admissions per Quarter",
       title = "(HB) Total Admissions across Health Boards in COVID Times per 
       Quarter, by Sex", 
       fill = "Sex")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```







<hr> 

## HEALTH BOARD - SIMD 

### HB - SIMD - Total Admissions 

```{r}
covid_admissions_HB_simd %>% 
  mutate(simd_quintile = fct_relevel(as.character(simd_quintile, 
                                  "1", "2", "3", "4", "5"))) %>%
  group_by(wk_ending_yr_month, simd_quintile) %>% 
  ggplot()+
  aes(x = number_admissions, 
      y = average20182019,
      colour = simd_quintile)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  labs(x = "Weekly Number of Admissions in COVID Times",
       y = "Average Weekly Number of Admissions in 2018/2019",
       title = "(HB) Weekly number of admissions in COVID Times against the 
       average of the same week in 2018/2019 by SIMD Levels",
      colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
  
```

So, creating monthly and quarterly total columns for each level of simd:
```{r}
covid_admissions_HB_simd <- covid_admissions_HB_simd %>% 
  group_by(simd_quintile, wk_ending_yr_month) %>% 
  mutate(total_admissions_per_month_simd = sum(number_admissions))

covid_admissions_HB_simd <- covid_admissions_HB_simd %>% 
  group_by(simd_quintile, wk_ending_yr_month) %>% 
  mutate(admissions_2018_19_per_month_simd = sum(average20182019))

covid_admissions_HB_simd <- covid_admissions_HB_simd %>% 
  group_by(simd_quintile, year_quarter) %>% 
  mutate(total_admissions_per_quarter_simd = sum(number_admissions))

covid_admissions_HB_simd <- covid_admissions_HB_simd %>% 
  group_by(simd_quintile, year_quarter) %>% 
  mutate(admissions_2018_19_per_quarter_simd = sum(average20182019))

covid_admissions_HB_simd
```

**PLOT total admissions per simd per month**
```{r}
covid_admissions_HB_simd %>% 
  mutate(simd_quintile = fct_relevel(as.character(simd_quintile, 
                                  "1", "2", "3", "4", "5"))) %>%
  group_by(wk_ending_yr_month, simd_quintile) %>% 
  ggplot()+
  aes(x = wk_ending_yr_month, 
      y = total_admissions_per_month_simd, 
      group = simd_quintile, 
      colour = simd_quintile)+
  scale_x_discrete(limits = dates) +
  geom_point()+
  geom_line()+
  labs(x = "Month and Year",
       y = "Total admissions",
       title = "(HB) Total admissions in COVID times per month by SIMD Level",
       subtitle = "January, 2020 - February 2022",
       colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

SIMD - monthly AVERAGE admissions 2018 and 2019 against COVID admissions 
```{r}
covid_admissions_HB_simd %>% 
   mutate(simd_quintile = fct_relevel(as.character(simd_quintile, 
                                  "1", "2", "3", "4", "5"))) %>%
  group_by(simd_quintile) %>% 
  ggplot()+
  aes(x = total_admissions_per_month_simd, 
      y = admissions_2018_19_per_month_simd,
      colour = simd_quintile)+
  geom_point()+
  labs(x = "Monthly number of admissions in COVID times",
       y = "Average monthly admissions in 2018-2019",
       title = "(HB) Monthly admissions in COVID times against the equivalent monthly 
       average admissions in previous years, by SIMD Level",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```  
BY QUARTER 
```{r}
covid_admissions_HB_simd %>% 
  mutate(simd_quintile = fct_relevel(as.character(simd_quintile, 
                                  "1", "2", "3", "4", "5"))) %>%
  ggplot()+
  aes(x = year_quarter, 
      y = total_admissions_per_quarter_simd,
      group = simd_quintile, 
      colour = simd_quintile)+
  geom_point()+
  geom_line()+
  labs(x = "Quarter and Year",
       y = "Total Admissions per Quarter",
       title = "(HB) Total Admissions across Health Boards in COVID Times per 
       Quarter, by SIMD", 
       colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

SIMD - quarterly AVERAGE admissions 2018 and 2019 against COVID admissions 
```{r}
covid_admissions_HB_simd %>% 
   mutate(simd_quintile = fct_relevel(as.character(simd_quintile, 
                                  "1", "2", "3", "4", "5"))) %>%
  group_by(simd_quintile) %>% 
  ggplot()+
  aes(x = total_admissions_per_quarter_simd, 
      y = admissions_2018_19_per_quarter_simd,
      colour = simd_quintile)+
  geom_point()+
  labs(x = "Quarterly number of admissions in COVID times",
       y = "Average quarterly admissions in 2018-2019",
       title = "(HB) Quarterly admissions in COVID times against the equivalent quarterly 
       average admissions in previous years, by SIMD Level",
       subtitle = "COVID: January, 2020 - February, 2022 / 
       Previous years: 2018 - 2019",
       colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```  



# STATISTICS SIMD COVID TIMES 

* Looking at COVID times - emergency admissions - by SIMD

I need the mean of total admissions for each simd quintile:
```{r}
covid_admissions_HB_simd %>% 
  group_by(simd_quintile) %>% 
  summarise(mean_total_admissions = mean(total_admissions_per_month_simd))

```


## SIMD 1 (most deprived) and SIMD 3 (mid deprivation)

$$H0 : μ_{SIMD\ 1:\ (most\ deprived)} - μ_{SIMD\ 3\ (mid\ deprived)} = 0$$
$$Ha : μ_{SIMD\ 1:\ (most\ deprived)} - μ_{SIMD\ 3\ (mid\ deprived)} > 0$$

```{r}
library(infer)
```

```{r}
covid_admissions_HB_simd %>% 
  ggplot(aes(x = simd_quintile, y = total_admissions_per_month_simd, group = simd_quintile))+
  geom_boxplot() +
  labs(x = "SIMD Quintile
       1 = Most Deprived : 5 = Least Deprived", 
       y = "Total admissions per month")

```

permutation to generate the null distribution
NOTE - we only want 1 and 3 just now
```{r}
covid_admissions_HB_simd_one_three <- covid_admissions_HB_simd %>% 
  filter(simd_quintile == "1" | simd_quintile == "3") %>% 
  mutate(simd_quintile = as.character(simd_quintile)) 

null_distribution_most_mid <- covid_admissions_HB_simd_one_three %>% 
  specify(total_admissions_per_month_simd~simd_quintile) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("1", "3"))
head(null_distribution_most_mid)

```

observed stat:
```{r}
observed_stat <- covid_admissions_HB_simd_one_three %>% 
  specify(total_admissions_per_month_simd~simd_quintile) %>% 
  calculate(stat = "diff in means", order = c("1", "3"))
observed_stat
```

```{r}
null_distribution_most_mid %>% 
  visualise()+
  shade_p_value(obs_stat = observed_stat, direction = "right")
```

```{r}
p_value <- null_distribution_most_mid %>% 
  get_p_value(obs_stat = observed_stat, direction = "right")
p_value

format.pval(p_value)
```

p = 2.22e-16 So very close to 0. 
Therefore as p < 0.05 we can reject our null hypothesis in favour of the 
alternative hypothesis.
The mean for total admissions for SIMD 1 (Most Deprived) is significantly 
greater than that of SIMD 3 (mid level deprivation area) so sugests that those
in more deprived areas are more likely to be admitted to hospital in COVID 
months. 

<br>
<hr>
<br>

# BETWEEN 3 and 5 (least deprived)


$$H0 : μ_{SIMD\ 3:\ (mid\ deprived)} - μ_{SIMD\ 5\ (least\ deprived)} = 0$$
$$Ha : μ_{SIMD\ 3:\ (mid\ deprived)} - μ_{SIMD\ 5\ (least\ deprived)} > 0$$


```{r}
covid_admissions_HB_simd %>% 
  ggplot(aes(x = simd_quintile, y = total_admissions_per_month_simd, group = simd_quintile))+
  geom_boxplot() +
  labs(x = "SIMD Quintile
       1 = Most Deprived : 5 = Least Deprived", 
       y = "Total admissions per month")

```

permutation to generate the null distribution
NOTE - we only want 3 and 5 just now
```{r}
covid_admissions_HB_simd_three_five <- covid_admissions_HB_simd %>% 
  filter(simd_quintile == "3" | simd_quintile == "5") %>% 
  mutate(simd_quintile = as.character(simd_quintile)) 

null_distribution_mid_least <- covid_admissions_HB_simd_three_five %>% 
  specify(total_admissions_per_month_simd ~ simd_quintile) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("3", "5"))
head(null_distribution_mid_least)

```

observed stat:
```{r}
observed_stat_mid_least <- covid_admissions_HB_simd_three_five %>% 
  specify(total_admissions_per_month_simd ~ simd_quintile) %>% 
  calculate(stat = "diff in means", order = c("3", "5"))
observed_stat_mid_least
```

```{r}
null_distribution_mid_least %>% 
  visualise()+
  shade_p_value(obs_stat = observed_stat_mid_least, direction = "right")
```


```{r}
p_value_mid_least <- null_distribution_mid_least %>% 
  get_p_value(obs_stat = observed_stat_mid_least, direction = "right")
p_value_mid_least

format.pval(p_value_mid_least)
```

p < 2.22e-16 So very close to 0. 
Therefore as p < 0.05 we can reject our null hypothesis in favour of the 
alternative hypothesis.
The mean for total admissions for SIMD 3 (Mid deprivation) is significantly 
greater than that of SIMD 5 (Least deprivation) so again suggests that those
in more deprived areas are more likely to be admitted to hospital in COVID 
months.

<br>
<hr>
<br>

Total monthly admissions for SIMD - measures of spread 

```{r}
covid_admissions_HB_simd %>%
  group_by(simd_quintile) %>%
  summarise(
    Mean = mean(total_admissions_per_month_simd),
    Q1 = quantile(total_admissions_per_month_simd, 0.25),
    Q2 = quantile(total_admissions_per_month_simd, 0.5),
    Q3 = quantile(total_admissions_per_month_simd, 0.75), 
    IQR = Q3 - Q1
  )
```

