---
title: "ae_waiting_time"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(e1071)
library(infer)
library(tidyverse)
library(janitor)
library(lubridate)
library(stringr)

```

```{r}

ae_waiting_time <- read_csv(here::here("data/ae_activity_waiting_times_apr2022-data.csv")) %>% 
  clean_names()

ae_waiting_time_prior <- read_csv(here::here("data/monthly_ae_waitingtimes_A&E attendances and performance data_202201.csv")) %>% 
  clean_names()

```

__Data Cleaning__

```{r}

ae_waiting_time_prior %>% 
  left_join(ae_waiting_time_prior, location_code, by = "treatment_location", all.x = TRUE)

ae_waiting_time_prior <- ae_waiting_time_prior %>% 
  mutate(year = str_c(
    str_sub(month, start = 1, end = 4)), .before = 1) %>% 
  mutate(month = str_c(
    str_sub(month, start = 5, end = 6)
  ))


ae_waiting_time_prior <- ae_waiting_time_prior %>% 
  filter(year >= 2016)

colnames(ae_waiting_time_prior)

ae_waiting_time_subset <- ae_waiting_time_prior %>% 
  select("year", "month" , "treatment_location" , "department_type", "number_of_attendances_aggregate",  "number_meeting_target_aggregate","attendance_greater8hrs", "attendance_greater12hrs", "discharge_destination_admission_to_same" , "discharge_destination_other_specialty" ,"discharge_destination_residence", "discharge_destination_transfer" ,"discharge_destination_unknown")

ae_waiting_time_subset <- ae_waiting_time_subset %>% 
  filter( year != "2016" | month != "01") %>%
  filter( year != "2016" | month != "02") %>%
  filter( year != "2016" | month != "03") %>%
  filter( year != "2016" | month != "04") %>%
  filter( year != "2016" | month != "05") %>%
  filter( year != "2016" | month != "06")

ae_waiting_time_subset <- ae_waiting_time_subset %>% 
  mutate(year_quarter = case_when(
    str_detect(month, "01|02|03") ~ "Q1",
    str_detect(month, "04|05|06") ~ "Q2",
    str_detect(month, "07|08|09") ~ "Q3",
    str_detect(month, "10|11|12") ~ "Q4"
  ), .after = 1)
           
ae_waiting_time_subset <- ae_waiting_time_subset %>% 
  mutate(year_quarter = str_c(year, year_quarter))

ae_waiting_time_clean <- ae_waiting_time_subset %>%   
   select(c(1:9))                                     

# filter(year != "2022") %>% 
# filter(year_quarter != "2021Q4") %>% 

ae_waiting_time_clean <- ae_waiting_time_clean %>% 
  mutate(number_meeting_target_aggregate = coalesce(number_meeting_target_aggregate, 0)) %>% 
  mutate(attendance_greater8hrs = coalesce(attendance_greater8hrs, 0)) %>% 
  mutate(attendance_greater12hrs = coalesce(attendance_greater12hrs, 0))

write_csv(ae_waiting_time_subset, file = here::here("data/ae_waiting_time_subset.csv"))

write_csv(ae_waiting_time_clean, file = here::here("data/ae_waiting_time_clean.csv"))

```

```{r}
library(ggplot2)
```


```{r}

ae_waiting_time_clean <- ae_waiting_time_clean %>% 
  mutate(year_quarter = as_factor(year_quarter)) 

```

```{r}
# avg_ae_waiting_time <- ae_waiting_time_clean %>% 
    group_by(year_quarter) %>% 
    summarise(mean(number_of_attendances_aggregate),
              mean(number_meeting_target_aggregate),
              mean(attendance_greater8hrs),
              mean(attendance_greater12hrs))
```


```{r}

ae_waiting_time_clean_long <- ae_waiting_time_clean %>% 
  rename("hrs_waiting_4" = number_meeting_target_aggregate,
         "hrs_more_waiting_8" = attendance_greater8hrs,
         "hrs_more_waiting_12" = attendance_greater12hrs) 

ae_waiting_time_clean_long_p <-ae_waiting_time_clean_long %>%
  pivot_longer(cols = c(hrs_waiting_4,hrs_more_waiting_8, hrs_more_waiting_12), names_to =  "hrs_waiting", values_to = "number_attendances") %>%
  mutate( year = as.numeric(year))

head(ae_waiting_time_clean_long_p)



```


__Merge Data__

```{r}
location_code <- ae_waiting_time %>% 
  select(3:5) %>% 
  rename("treatment_location" = "treatment_location_code")
```

```{r}

ae_waiting_time_clean_long_p <- left_join(ae_waiting_time_clean_long_p, location_code, by = "treatment_location") %>% 
  unique() %>% 
  relocate(treatment_location_name, .after = 4) %>% 
  relocate(hb_treatment_description, .after = 4)
  

ae_waiting_time_clean_long_p <- ae_waiting_time_clean_long_p %>% 
  filter(department_type == "Emergency Department") %>% 
  rename(region = hb_treatment_description)

write_csv(ae_waiting_time_clean_long_p, file = here::here("data/ae_waiting_time_clean_long_p.csv"))

```

__Plots__
Evaluate the number of attendances for each range waiting time

```{r}

#Create tibble to join as a large dataset

 plot_data_4hrs <- ae_waiting_time_clean_long_p %>% 
  group_by(year_quarter) %>% 
  filter(hrs_waiting == "hrs_waiting_4") %>% 
  summarise(total_attendances = sum(number_attendances)) %>% 
  mutate(time_waiting = "4hrs", .after = 1)

plot_data_8hrs <- ae_waiting_time_clean_long_p %>% 
  group_by(year_quarter) %>% 
  filter(hrs_waiting == "hrs_more_waiting_8") %>% 
  summarise(total_attendances = sum(number_attendances)) %>% 
  mutate(time_waiting = "8hrs", .after = 1)

plot_data_12hrs <- ae_waiting_time_clean_long_p %>% 
  group_by(year_quarter) %>% 
  filter(hrs_waiting == "hrs_more_waiting_12") %>% 
  summarise(total_attendances = sum(number_attendances)) %>% 
  mutate(time_waiting = "12hrs", .after = 1)

waiting_time_all_range <- rbind(plot_data_4hrs, plot_data_8hrs, plot_data_12hrs)

write_csv(waiting_time_all_range, file = here::here("data/waiting_time_all_range.csv"))

#Number of People in Waiting Time Range
  waiting_time_all_range %>% 
  group_by(year_quarter) %>% 
  ggplot()+
  aes(x = year_quarter, y = log10(total_attendances), group = time_waiting, colour = time_waiting)+
  geom_point()+
  geom_line()+
  xlab("Year(Quarter)")+ylab("Number of Attendances (log 10)")+
  theme(axis.text.x = element_text(angle = 90, size = 5, vjust = 0.5, hjust=1))+
  labs(color="Time Waiting", size = 5)+
  ggtitle("Number of People in Waiting Time Range")

  
```
_Insights_


__ready to go__
The total number of attendances from 2016 to 2022
```{r}

#Number of A&E Attendances from 2016 to 2022
ae_waiting_time_clean_long_p %>% 
  group_by(year_quarter) %>% 
  summarise(total_attendances = sum(number_attendances)) %>% 
  ggplot()+
  aes(x = year_quarter, y = log10(total_attendances))+
  geom_point()+
  geom_line(group = 1)+
  xlab("Year(Quarter)")+ylab("Number of Attendances(log 10)")+
  theme(axis.text.x = element_text(angle = 90, size = 5, vjust = 0.5, hjust=1))+
  labs(color="Time Waiting", size = 5)+
  ggtitle("Number of A&E Attendances from 2016 to 2022")


```


__Statistics__




In this case we are going to check the difference between mean attendances before Covid (~2019Q4) and after (2020Q1~)

Hypothesis : two mean independent sample test with permutation.

H0: after_Covid_attendances_mean = before_Covid_attendances_mean
Ha: after_Covid_attendances_mean != before_Covid_attendances_mean

```{r}
#before Covid attendances mean

before_covid_attendances_avg <- ae_waiting_time_clean%>% 
  group_by(year_quarter) %>% 
  filter(year != "2022") %>% 
  filter(year != "2021") %>% 
  filter(year_quarter != "2019Q4") %>% 
  summarise(avg_attendances = mean(number_of_attendances_aggregate)) %>% 
  mutate(status = "before")
  
after_covid_attendances_avg <- ae_waiting_time_clean%>% 
  #group_by(year_quarter ) %>% 
  filter(year != "2016") %>%
  filter(year != "2017") %>% 
  filter(year != "2018") %>% 
  filter(year_quarter != "2019Q1") %>% 
  filter(year_quarter != "2019Q2") %>% 
  filter(year_quarter != "2019Q3") %>% 
  summarise(avg_attendances= mean(number_of_attendances_aggregate)) %>% 
  mutate(status = "after")

waiting_time_mean <- rbind(before_covid_attendances_avg, after_covid_attendances_avg)


waiting_time_mean %>% 
  ggplot(aes(x = status, y = avg_attendances))+
  geom_boxplot()+
  xlab("After vs Before Covid")+ylab("Avg Attendances (quater)")


```

The diagram indicates that distribution of the average attendances before Covid (2019Q3) is at around 1500; while Covid and after is at around 1700 to 1800


__Try to do some stats__

```{r}

waiting_time_distribution <- waiting_time_mean %>% 
  specify(avg_attendances ~ status) %>% 
  hypothesize(null = "independence") %>% 
  generate( reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("after", "before"))

observed_waiting_time <- waiting_time_mean %>% 
  specify(avg_attendances ~ status) %>% 
  calculate( stat = "diff in means", order = c("after", "before"))

waiting_time_distribution %>% 
  visualise()+
shade_p_value(obs_stat = observed_waiting_time, direction = "right")


```


```{r}

waiting_time_distribution %>% 
get_p_value(obs_stat = observed_waiting_time, direction = "right")

```


The H0 is rejected. There is no evidence proof that the mean attendances after Covid is the same as the mean attendances before Covid. Since this is a right-sided test, we can confirm that the mean attendances after Covid is greater than the mean attendance before Covid.


__explore the region__


```{r}

region_waiting_time <- ae_waiting_time_clean_long_p %>% 
  group_by(year_quarter, region) %>% 
  summarise(attendances = mean(number_of_attendances_aggregate))

region_waiting_time %>% 
  group_by(year_quarter) %>% 
  ggplot()+
  aes(x = year_quarter, y = attendances, group = region, color = region)+
  geom_line()+
  geom_point()+
  xlab("Year(Quarter)")+ylab("Number of Attendances")+
  theme(axis.text.x = element_text(angle = 90, size = 5, vjust = 0.5, hjust=1))+
  labs(color="Time Waiting", size = 3)+
  ggtitle("Average Number of Attendances (Region)")

region_waiting_time


write_csv(region_waiting_time, file = here::here("data/region_waiting_time.csv"))

```

__Ethical and legal considerations including data bias__

The datasets in this report is the open data authorized by NHS Scotland and is under NHS  Scotland 's protection. We do not own the copyright of the data. The material contained in this report is distributed without profit for research and educational purposes. 

The datasets recorded and collected the information of every episodes accessing the hospital services from the year of 2007 to 2022, which may not consent by patients. The data is anonymous. It contains without sensitive personal details such as address, occupancy and the reasons for accessing hospital services. However, there is a concern that the third party accessing the data for business usage. 

The data may have bias since it is not standardized some health board are not able to provide a full record of episodes due to the limitation of resources, which may cause difference to the result.
