---
title: "Demographic Analysis"
subtitle: rshiny group project
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(ggplot2)
```


```{r}
# hosp_activity_page <- read_csv(here("data/hospital_activity_page.csv"))
# hosp_activity_by_speciality <- 
#   read_csv(here("data/Hospital Activity by Speciality.csv"))

hosp_activity_and_demographics <- 
  read_csv(here("data/Hospital Activity and Patient Demographics.csv"))
hosp_activity_and_deprivation <- 
  read_csv(here("data/Hospital Activity and Deprivation.csv"))
# 
# aAndE_waiting_times <- 
#   read_excel(here("data/monthly_ae_waitingtimes_A&E attendances and performance data_202201.xlsx"))
```


```{r}
covid_ads_HB_deprivation <- 
  read_excel(here("data/Hospitalisations due to Covid 19/Admissions By Health Board and Deprivation_20220302.xlsx"))

# covid_ads_HB_and_speciality <- 
#   read_excel(here("data/Hospitalisations due to Covid 19/Admissions By Health Board and Specialty_20220302.xlsx"))
covid_ads_HB_age_sex <- 
  read_excel(here("data/Hospitalisations due to Covid 19/Admissions By Health Board, Age and Sex_20220302.xlsx"))
covid_ads_HSCP_deprivation <- 
   read_excel(here("data/Hospitalisations due to Covid 19/Admissions By HSCP and Deprivation_20220302.xlsx"))
# covid_ads_HSCP_speciality <- 
#   read_excel(here("data/Hospitalisations due to Covid 19/Admissions By HSCP and Specialty_20220302.xlsx"))
covid_ads_HSCP_age_sex <- 
  read_excel(here("data/Hospitalisations due to Covid 19/Admissions By HSCP, Age and Sex_20220302.xlsx"))
```


Clean names
```{r}
hosp_activity_and_demographics <- 
  janitor::clean_names(hosp_activity_and_demographics)
hosp_activity_and_deprivation <- 
  janitor::clean_names(hosp_activity_and_deprivation)
covid_ads_HB_deprivation <- 
  janitor::clean_names(covid_ads_HB_deprivation)
covid_ads_HB_age_sex <- 
  janitor::clean_names(covid_ads_HB_age_sex)
covid_ads_HSCP_deprivation <- 
  janitor::clean_names(covid_ads_HSCP_deprivation)
covid_ads_HSCP_age_sex <- 
  janitor::clean_names(covid_ads_HSCP_age_sex)
```



```{r}
names(hosp_activity_and_demographics)
names(hosp_activity_and_deprivation)
names(covid_ads_HB_deprivation)
names(covid_ads_HB_age_sex)
names(covid_ads_HSCP_deprivation)
names(covid_ads_HSCP_age_sex)
```

```{r}
hosp_activity_and_demographics %>% 
  distinct(age)
# 10 year increments i.e. 0-9, 10-19... 80-89, 90 years and over
# No NA values
hosp_activity_and_demographics %>% 
  mutate(age = is.na(age)) %>% 
  filter(age == TRUE)
```

```{r}
# 7 General hosp admission types for both demographics and deprivation:
# Elective Inpatients
# Emergency Inpatients,  (THIS IS THE ONLY DEFINITELY ACUTE ONE)
# Transfers
# All Day cases
# All Inpatients
# All inpatients and Day Cases
# Not Specified 
# No NAs.
hosp_activity_and_demographics %>% 
  distinct(admission_type)
hosp_activity_and_deprivation %>% 
  distinct(admission_type)

# 3 COVID ADMISSION TYPES for HB and HSCP - deprivation and age/sex: 
# All - Emergency - Planned
covid_ads_HB_deprivation %>% 
  distinct(admission_type)
covid_ads_HB_age_sex %>% 
  distinct(admission_type)
covid_ads_HSCP_deprivation %>% 
  distinct(admission_type)
covid_ads_HSCP_age_sex %>% 
  distinct(admission_type)  


# tried them all for NA and there are none
# hosp_activity_and_demographics %>% 
#   mutate(admission_type = is.na(admission_type)) %>% 
#   filter(admission_type == TRUE)
```

<br>

<hr>

## General Hospital Admissions 

### November 14th 2019 - February 22nd 2020

* **This data is by QUARTER (COVID admissions is by week ending from May 2020 to** 
**March 2022**
* **SEX here is "Male", "Female", COVID data also has Male, Female, and All**
* **AGE groups are in 10 year increments with (from "0-10 to "90 and over")**
  **COVID data is a lot dirtier**
* **These are general hospital admissions - not by health board etc**

<br>

#### Sex 

<br>
```{r}
# 2 sex = Female, Male 
hosp_activity_and_demographics %>% 
  distinct(sex)

hosp_activity_and_demographics %>% 
  mutate(sex = is.na(sex)) %>% 
  filter(sex == TRUE)
# No NAs
```
<br> 

**Sex against time and admissions**

Filter for acute services: 

```{r}
# might want to make use of this if we're not sure about admission types
# acute_target <- c("Emergency Inpatients", "Not Specified")
# Others = Elective Inpatients, Transfers, All Day cases, All Inpatients, All 
# Inpatients and Day cases, Not Specified

general_admissions_acute_age_sex <- hosp_activity_and_demographics %>% 
  filter(admission_type == "Emergency Inpatients")
general_admissions_acute_age_sex
```


Now group by and create graphs over time: 

```{r}

gen_admissions_sex_per_quarter <- general_admissions_acute_age_sex %>% 
  group_by(quarter, sex) %>% 
  summarise(total_admissions_per_quarter = n())
  
gen_admissions_sex_per_quarter %>% 
  ggplot()+
  aes(x = quarter, 
      y = total_admissions_per_quarter, 
      group = sex, colour = sex)+
  geom_line() + 
  labs(x = "Quarter", 
       y = "Total Admissions", 
       title = "Total Emergency Inpatient Admissions",
       subtitle = "Q3 2016 to Q3 2021", 
       colour = "Sex") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))

```

<br>

#### Average length of stay by sex - general admissions

```{r}

#this has negative numbers in it (?)
# general_admissions_acute_age_sex %>% 
#   drop_na(average_length_of_stay) %>% 
#   distinct(desc(average_length_of_stay))

# highlight_winter <- general_admissions_acute_age_sex %in% c("2016Q4", 
# "2017Q1", "2017Q4", "2018Q1", "2018Q4", "2019Q1", "2019Q4", "2020Q1", 
# "2020Q4", "2021Q1")
# 
# library(gghighlight)

#Might be best to split into quarters? Have one of those sliding scales for time
# Would also like to highlight the winter months 
general_admissions_acute_age_sex %>% 
  drop_na(average_length_of_stay) %>% 
  group_by(quarter, sex, average_length_of_stay) %>% 
  ggplot()+
  aes(x = quarter, 
      y = average_length_of_stay, 
      fill = sex)+
  geom_col(position = "dodge")+ 
  labs(x = "Quarter of each year", 
       y = "Average Length of stay (days)", 
       title = "Average Length of Stay for Emergency Inpatient Admissions By 
       Gender",
       subtitle = "Q3 2016 to Q3 2021", 
       colour = "Sex") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))

```




<br>

#### Age

```{r}

hosp_activity_and_demographics %>% 
  distinct(age) # 10 year increments i.e. 0-9, 10-19... 80-89, 90 years and over
# No NA values 

hosp_activity_and_demographics %>% 
  mutate(age = is.na(age)) %>% 
  filter(age == TRUE) # no NAs
```

```{r}
gen_admissions_age_per_quarter <- general_admissions_acute_age_sex %>% 
  # remember this is only looking at Emergency Inpatients but there are 7 groups
  group_by(quarter, age) %>% 
  summarise(total_admissions_per_quarter = n())
  
# Note that it's 0-9years that are the low values and I'll make this better soon
gen_admissions_age_per_quarter %>% 
  ggplot()+
  aes(x = quarter, 
      y = total_admissions_per_quarter, 
      group = age, colour = age)+
  geom_line() + 
  labs(x = "Yearly Quarter", 
       y = "Total Admissions", 
       title = "Total Emergency Inpatient Admissions",
       subtitle = "Q3 2016 to Q3 2021", 
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))

gen_admissions_age_per_quarter %>% 
  ggplot()+
  aes(x = quarter, 
      y = total_admissions_per_quarter, 
      group = age)+
  geom_col() + 
  facet_wrap(~age)+
  labs(x = "Yearly Quarter", 
       y = "Total Admissions", 
       title = "JUST MADE THIS TO SEE THE DIFFERENCE - AWARE IT'S MESSY - 
       Total Emergency Inpatient Admissions",
       subtitle = "Q3 2016 to Q3 2021") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))
```
<br> 
#### average length of stay - general admissions - age 

```{r}

# as before, this has negative numbers in it (?)
# general_admissions_acute_age_sex %>% 
#   drop_na(average_length_of_stay) %>% 
#   distinct(desc(average_length_of_stay))


#Might be best to split into years?
# Have one of those sliding scales for time
# Would also like to highlight the winter months 
general_admissions_acute_age_sex %>% 
  drop_na(average_length_of_stay) %>% 
  group_by(quarter, age, average_length_of_stay) %>% 
  ggplot()+
  aes(x = quarter, 
      y = average_length_of_stay, 
      fill = age)+
  geom_col(position = "dodge") +  
  facet_wrap(~age)+
  labs(x = "Quarter of each year", 
       y = "Average Length of stay (days)", 
       title = "Average Length of Stay for Emergency Inpatient Admissions By 
       Age Group",
       subtitle = "Q3 2016 to Q3 2021", 
       colour = "Age Group") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))

```


<hr> 
#### SIMD (Scottish Index of Multiple Deprivation)

* This is a quintile scale (1 = "Most Deprived", 5 = "Least Deprived")
* The most appropriate SIMD released for each year is used. 

<br>

```{r}
# We have the same admission types here as above so I'll do the same - i.e. only
# include the Emergency Inpatients though there are 7 groups
hosp_activity_and_deprivation %>% 
  distinct(admission_type)


general_admissions_acute_deprivation <- hosp_activity_and_deprivation %>% 
  filter(admission_type == "Emergency Inpatients")
general_admissions_acute_deprivation

```

SIMD Types
* note there are 962 NAs: there are codes in the simdqf column as to why 
* sometimes it's just a general geographical group 
```{r}
general_admissions_acute_deprivation %>% 
  group_by(simd) %>% 
  summarise(total_for_each_simd = n())
```
```{r}
gen_admissions_simd_per_quarter <- general_admissions_acute_deprivation %>% 
  group_by(quarter, simd) %>% 
  drop_na(simd) %>% 
  summarise(total_admission_per_quarter = n())
gen_admissions_simd_per_quarter


gen_admissions_simd_per_quarter %>% 
  ggplot() + 
  aes(x = quarter, 
      y = total_admission_per_quarter, 
      fill = simd) + 
  geom_col(position = "dodge") + #why won' it dodge? Not enough space??
  labs(x = "Yearly Quarter", 
       y = "Total Admissions", 
       title = "Total Emergency Inpatient Admissions by Deprivation Level")+#, 
       #colour = "SIMD Level: (1 = Most Deprived : 5 = Least Deprived")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))

gen_admissions_simd_per_quarter %>% 
  ggplot() + 
  aes(x = quarter, 
      y = total_admission_per_quarter, 
      fill = simd) + 
  geom_col(position = "dodge") + 
  facet_wrap(~simd)+
  labs(x = "Yearly Quarter", 
       y = "Total Admissions", 
       title = "Total Emergency Inpatient Admissions by Deprivation Level", 
       colour = "SIMD Level: (1 = Most Deprived : 5 = Least Deprived")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))


# hmmm, that's not a very nice line graph:
gen_admissions_simd_per_quarter %>% 
  ggplot() + 
  aes(x = quarter, 
      y = total_admission_per_quarter, 
      group = simd, 
      colour = simd) + 
  geom_line() +
  labs(x = "Yearly Quarter", 
       y = "Total Admissions", 
       title = "Total Emergency Inpatient Admissions by Deprivation Level")+#, 
       #colour = "SIMD Level: (1 = Most Deprived : 5 = Least Deprived")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```





<hr>
<br>
## COVID ADMISSIONS 

### May 21st 2020 - March 2nd, 2022

#### Sex and gender by healthboard

<br>

NOTE 
sex: "All" = 33951 rows, Female = 4782, Male = 4783 

<br> 

```{r}
# by healthboard

# admission types are either emergncy or planned - that's good news for us
covid_ads_HB_age_sex %>% 
  distinct(admission_type)

covid_acute_HB_age_sex <- covid_ads_HB_age_sex %>% 
  filter(admission_type == "Emergency")

# this is not neatly divided and there is an age group "41760" hmm
covid_acute_HB_age_sex %>% 
  group_by(age_group) %>% 
  summarise(count_in_age_group = n())
```

<br>

```{r}
# nope
covid_acute_HB_age_sex %>% 
  group_by(age_group) %>% 
  ggplot()+
  aes(x = number_admissions, 
      y = average20182019, 
      fill = age_group)+
  geom_point()
  
```


```{r}
covid_ads_HB_age_sex %>% 
  group_by(sex) %>% 
  summarise(total_of_each_sex = n())

covid_ads_HB_demographics_sex <- covid_ads_HB_age_sex %>% 
  group_by(week_ending, sex) %>% 
  summarise(total_admissions_for_week_ending = n())
covid_ads_HB_demographics_sex


covid_ads_HB_demographics_sex %>% 
  ggplot()+
  aes(x = week_ending, 
      y = total_admissions_for_week_ending, 
      group = sex, 
      colour = sex)+
  geom_line()+
  labs(title = "I AM AWARE THIS IS AWFUL")
```

