---
title: "Demographics analysis 1 - General Admissions"
subtitle: General Admissions
output:
  html_document:
    df_print: paged
---

* **These are general hospital admissions - 7 admission_types (1 is acute)**
* **This data is by QUARTER - 2016Q3 to 2021Q3 - (COVID admissions is by "week** 
**ending" from May 2020 to March 2022** 
* **SEX here is "Male", "Female", COVID data also has Male, Female, and All**
* **AGE groups are in 10 year increments with (from "0-10 to "90 and over")**
  **COVID data is a lot dirtier**
* Total per age group 
  * 0-9 years	1261	
  * 10-19 years	1893	
  * 20-29 years	1965	
  * 30-39 years	1990	
  * 40-49 years	2041	
  * 50-59 years	2091	
  * 60-69 years	2114	
  * 70-79 years	2129	
  * 80-89 years	2140	
  * 90 years and over	2007


## Libraries Used 

```{r}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(ggplot2)
library(lubridate)
library(stringr)
```
<br> 

## Read In Data

```{r}
hosp_activity_agesex <- read_csv(here("../../raw_data/inpatient_and_daycase_by_nhs_board_of_treatment_age_and_sex.csv"))

hosp_activity_simd <- 
  read_csv(here("../../raw_data/inpatient_and_daycase_by_nhs_board_of_treatment_and_simd.csv"))
```
<br>

## Clean Data

```{r}
hosp_activity_agesex <- janitor::clean_names(hosp_activity_agesex)
# glimpse(hosp_activity_agesex)

hosp_activity_simd <- janitor::clean_names(hosp_activity_simd)
# glimpse(hosp_acivity_simd)
```
 
<br>

We only need ACUTE patients, what do we have:

admission_type:
* Elective Inpatients		(not acute)
* *Emergency Inpatients*				
* Transfers			(not acute)			
* All Day cases				(not acute)		
* All Inpatients		  		(not necasseraly acute)		
* All Inpatients and Day cases			(not acute)			
* Not Specified   		(not acute???)
```{r}
hosp_activity_agesex %>% 
  group_by(admission_type) %>% 
  summarise(total = n())

hosp_activity_simd %>% 
  group_by(admission_type) %>% 
  summarise(total = n())
```
<br>

removing those that aren't acute (see above)

```{r}
hosp_acute_activity_agesex <- hosp_activity_agesex %>% 
  filter(admission_type == "Emergency Inpatients")

hosp_acute_activity_simd <- hosp_activity_simd %>% 
  filter(admission_type == "Emergency Inpatients")

```


### Seeing what we're working with:

*Age*

```{r}
hosp_acute_activity_agesex %>% 
  group_by(age) %>% 
  summarise(total_patients_in_age_group = n())

# 10 year increments i.e. 0-9, 10-19... 80-89, 90 years and over
# No NA values

hosp_acute_activity_agesex %>% 
  mutate(age = is.na(age)) %>% 
  filter(age == TRUE)
```

*Sex*

```{r}
hosp_acute_activity_agesex %>% 
  group_by(sex) %>% 
  summarise(total_patients_in_sex_group = n())

# Female: 9,832   Male: 9,799 
# No NA values

hosp_acute_activity_agesex %>% 
  mutate(sex = is.na(age)) %>% 
  filter(sex == TRUE)
```

**SIMD** 
```{r}
hosp_acute_activity_simd %>% 
  group_by(simd) %>% 
  summarise(total = n())
#There are 962 NAs and ~ 1000 of levels 1:5... remember to tidy this in analysis
```

<br>

```{r}
quarter_dates <- c("2016Q3", "2016Q4", "2017Q1", "2017Q2", "2017Q3", "2017Q4", 
                   "2018Q1", "2018Q2", "2018Q3", "2018Q4", "2019Q1", "2019Q2", 
                   "2019Q3", "2019Q4", "2020Q1", "2020Q2", "2020Q3", "2020Q4",
                   "2021Q1", "2021Q2", "2021Q3")
```

<hr>

<br>

## Graphs

Remember there were 7 admission types... I chose to keep one. 

Might want to make use of this if we're not sure about admission types:
 `acute_target <- c("Emergency Inpatients", "Not Specified")`

Others = Elective Inpatients, Transfers, All Day cases, All Inpatients, All 
Inpatients and Day cases, Not Specified

<br>

### Graphs of general admissions and sex

#### Sex against time and admissions

```{r}
# do I want to use total length of stay or total average_length_of_stay
# need a total column: 
# sometimes stay = 0, and length of stay = 5

hosp_acute_activity_agesex <- hosp_acute_activity_agesex %>% 
  group_by(sex, quarter) %>% 
  mutate(total_stays_per_quarter_sex = sum(stays))
hosp_acute_activity_agesex
```


ALSO WANT TOTALS FOR AGE BY QUARTER 
```{r}

hosp_acute_activity_agesex <- hosp_acute_activity_agesex %>% 
  group_by(age, quarter) %>% 
  mutate(total_stays_per_quarter_age = sum(stays))
hosp_acute_activity_agesex
```





## Sex and stays

```{r}
hosp_acute_activity_agesex %>% 
  ggplot()+
  aes(x = quarter, 
      y = total_stays_per_quarter_sex, 
      group = sex, colour = sex) +
  geom_line() + 
  labs(x = "Yearly Quarter", 
       y = "Total Stays", 
       title = "Total Emergency Inpatient Stays Across Scotland by Sex",
       subtitle = "Q3, 2016 - Q3, 2021", 
       colour = "Sex") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))

```


<br>

## AGE GROUPS AND TOTAL STAYS 

```{r}
hosp_acute_activity_agesex %>% 
  ggplot()+
  aes(x = quarter, 
      y = total_stays_per_quarter_age, 
      group = age, colour = age) +
  geom_point()+
  geom_line() + 
  labs(x = "Yearly Quarter", 
       y = "Total Stays", 
       title = "Total Emergency Inpatient Stays Across Scotland by Age Group",
       subtitle = "Q3, 2016 - Q3, 2021", 
       colour = "Age") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))

```

## SIMD - deprivation level data 


SIMD (**Scottish Index of Multiple Deprivation**)

* This is a quintile scale (1 = "Most Deprived", 5 = "Least Deprived").
* The most appropriate SIMD released for each year is used. 
* REMEMBER: In the simd level column, there are 962 NAs and ~1000 of levels 
  1:5. Remember to tidy this in analysis. There are codes in the simdqf column 
  as to why they are NA - sometimes it's just a general geographical groupings. 

```{r}
hosp_acute_activity_simd %>% 
  group_by(simd) %>% 
  summarise(total = n())
#There are 962 NAs and ~ 1000 of levels 1:5... remember to tidy this in analysis
```



<br>


Drop NAs
```{r}
hosp_acute_activity_simd_no_na <- hosp_acute_activity_simd %>% 
  drop_na(simd)
hosp_acute_activity_simd_no_na
```

```{r}

hosp_acute_activity_simd_no_na_simd <- hosp_acute_activity_simd_no_na %>% 
  group_by(simd, quarter) %>% 
  mutate(total_stays_per_quarter_simd = sum(stays))
hosp_acute_activity_simd_no_na_simd

```


```{r}
hosp_acute_activity_simd_no_na_simd %>%
     mutate(simd = fct_relevel(as.character(simd, 
                                  "1", "2", "3", "4", "5"))) %>%
  ggplot()+
  aes(x = quarter, 
      y = total_stays_per_quarter_simd, 
      group = simd, colour = simd) +
  geom_point()+
  geom_line() + 
  labs(x = "Yearly Quarter", 
       y = "Total Stays", 
       title = "Total Emergency Inpatient Stays Across Scotland by SIMD Level",
       subtitle = "Q3, 2016 - Q3, 2021", 
       colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))

```



## EPISODES 

```{r}

hosp_acute_activity_simd_no_na_simd <- hosp_acute_activity_simd_no_na %>% 
  group_by(simd, quarter) %>% 
  mutate(total_episodes_per_quarter_simd = sum(episodes))
hosp_acute_activity_simd_no_na_simd

```


```{r}
hosp_acute_activity_simd_no_na_simd %>%
     mutate(simd = fct_relevel(as.character(simd, 
                                  "1", "2", "3", "4", "5"))) %>%
  ggplot()+
  aes(x = quarter, 
      y = total_episodes_per_quarter_simd, 
      group = simd, colour = simd) +
  geom_point()+
  geom_line() + 
  labs(x = "Yearly Quarter", 
       y = "Total Stays", 
       title = "Total Emergency Inpatient Episodes Across Scotland by SIMD Level",
       subtitle = "Q3, 2016 - Q3, 2021", 
       colour = "SIMD Level:
       1 = Most Deprived
       5 = Least Deprived") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=0.9))

```
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>
<hr>


